How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in the Pacific Pond Turtle (\textit{Emys marmorata})

### 2015-09-08 doc/draft_turtle.tex
abstract, jim ken comments, need new graphs

\begin{abstract}
  ABSTRACT NEEDS TO BE WRITTEN
\end{abstract}

@@

% cryptic diversity
%   most species are still deliminated solely based on morphology
Molecular systematics has repeatedly demonstrated the existence of cryptic species that can only be diagnosed using genetic data \citep{Stuart2006,Bickford2007,SchlickSteiner2007,Pfenninger2007,Clare2011,Funk2012}. In attempts to streamline the documentation of biodiversity, several methods of species delimitation that rely almost entirely on genetic data have recently been proposed \citep{Yang2010b}. Although strong caveats on the utility of these methods have been raised \citep{Bauer2000,Carstens2013}, they are nevertheless being used to name species \citep{Leache2010,Spinks2014}.

@@

These considerations have sparked interest in whether geometric morphometric analyses can capture fine-scale variation that can be used for identifying cryptic species. Most such studies focus on using morphometrics to discover differences between taxa that were identified by other means \citep{Polly2003,Zelditch2004,Gaubert2005b,Gunduz2007,Polly2007a,Demandt2009}. Additionally, there has been a fair amount of work on automated taxon identification and classification of taxa into groups \citep{Baylac2003,Dobigny2003,MacLeod2007,VandenBrink2011}. In cases where genetic data are ambiguous or lacking for many samples, morphometric approaches could help identify cryptic species. This would make the task of identifying and maintaining endangered or conserved groups much easier and could contribute to improved classifications of extinct taxa and populations.

@@

% past work on automatic taxon identification and older approaches to classifying taxa
% why use machine learning methods
\subsection{Background and study system}
Machine learning is an extension of known statistical methodology \citep{Hastie2009} that emphasizes high predictive accuracy and generality at the expense of the interpretability of individual parameters. The basic statistical mechanics are supplemented by randomization, sorting, and partitioning algorithms, along with the maximization or minimization of summary statistics, in order to best estimate a general model for all data, both sampled and unsampled \citep{Hastie2009}. Machine learning approaches have found use in medical research, epidemiology, economics and automated image identification such as handwritten zip codes \citep{Hastie2009}. The two major classes of machine learning methods are unsupervised and supervised learning. Unsupervised learning methods are used with unlabeled data where the underlying structure is estimated, and they are analogous to clustering and density estimation methods \citep{Kaufman1990}. Supervised learning methods are used with labeled data where the final output of data is known and the rules for going from input to output are inferred. These are analogous to classification and regression models \citep{Breiman1984}. Out application of the alternative approaches used in this study illustrates only a sampling of the various previously-derived methods for clustering observations and fitting classification models.

@@

Here, we used multiple machine learning methods, both unsupervised and supervised, to compare different classification hypotheses. These methods provide different and unique advantages for understanding how to classify taxa, as well as the accuracy of the resulting classifications. Although machine learning methods such as neural networks have been applied to studying shape variation \citep{Baylac2003,Dobigny2003,MacLeod2007,VandenBrink2011}, including in the context of automated taxon identification and classification of groups, the number of cases remains limited. In the current study, we not only consider pure classification accuracy but also use a statistic of classification strength that reflects the rate at which taxa are both accurately and inaccurately classified.

We analyzed the problem of whether there are distinct subspecies or cryptic species within the western pond turtle, \textit{Emys marmorata} BAIRD CITATIONS (formerly \emph{Clemmys marmorata}; see \citealp{Feldman2002}). \textit{Emys marmorata} is distributed from northern Washington State, USA to Baja California, Mexico. Traditionally, \textit{E. marmorata} was classified into two named subspecies: the northern \textit{E. marmorata marmorata} and the southern \textit{Emys marmorata pallida} \citep{Seeliger1945}, with a central Californian intergrade zone in between. \textit{Emys marmorata marmorata} is differentiated from \textit{E. marmorata pallida} by the presence of a pair of triangular inguinal scales and darker neck markings. The triangular inguinal plates can sometimes be present in \textit{E. marmorata pallida} though they are considerably smaller. \citet{Seeliger1945} did not formally include the Baja California populations of \textit{E. marmorata} in either taxon, implying the existence of a third distinct but unnamed subspecies.

Previous work on morphological variation in \textit{E. marmorata} has focused primarily on differentiation between populations over a portion of the species' total range \citep{Lubcke2007,Germano2008,Germano2009,Bury2010}; comparatively few studies have included specimens from across the entire range \citep{Holland1992}. Most of these studies considered how local biotic and abiotic factors may contribute to differences in carapace length and found that size can vary greatly between different populations \citep{Lubcke2007,Germano2008,Germano2009}. There also has been interest in size-based sexual dimorphism in \textit{E. marmorata} \citep{Holland1992,Lubcke2007,Germano2009}, with males being on average larger than females based on total carapace length and other linear measurements. However, the quality of size as a classifier of sex can vary greatly between populations \citep{Holland1992} because of the amount of size differences among populations \citep{Lubcke2007,Germano2009}. However, the effect of sexual dimorphism on shape, \textit{sensu} \citet{Kendall1977a}, has not been assessed \citep{Holland1992,Lubcke2007,Germano2008}.

@@

More recently, the phylogeography of \textit{E. marmorata} and the possibility of cryptic diversity was investigated using molecular data \citep{Spinks2005,Spinks2010,Spinks2014}. Based on mitochondrial DNA, \citet{Spinks2005} recognized four subclades within \textit{E. marmorata}, a northern clade, a San Joaquin Valley clade, a Santa Barbara clade, and a southern clade. Analyses with nuclear DNA \citep{Spinks2010} with single-nucleotide polymorphism (SNP) data suggest a primarily north--south division in \textit{E. marmorata}, although the dataset differed from that of \citet{Spinks2005} in the location of this break point. These studies discussed the potential taxonomic implications of their results, with \citet{Spinks2014} going so far as to strongly advocate for the recognition of at least two species (\emph{E. marmorata} and \emph{E. pallida}), and a possible third based on populations in Baja California. However, they did not discuss in detail the morphological characters that would help to diagnose these species beyond those specified by \citet{Seeliger1945}. Given that these characters are somewhat variable within the proposed species, and that \citet{Holland1992} described shell shape variation that might be consistent with this taxonomy, a geometric morphometric analysis of shell shape might provide a reliable way to diagnose groups (whether species or subspecies) within \textit{E. marmorata}.

@@

\section{Materials and Methods}
\subsection{Specimens, sampling, morphometrics}
We collected landmark-based morphometric data from 354 adult \textit{E. marmorata} museum specimens. These specimens are a subset of those included in \citet{Angielczyk2007}, \citet{Angielczyk2011}, and \citet{Angielczyk2013a}, essentially representing the adult individuals included in those papers. We chose to focus on adults because significant changes in plastron shape occur over the course of ontogeny in \textit{E. mamorata} and other emydines \citep{Angielczyk2013a}. \uppercase{Peter note -- how was something classified as an adult?}

We assigned a classification to each specimen for the different binning schemes based on geographic occurrence data recorded in museum collection archives. When precise latitude and longitude information were not available we estimated them from locality information. Because \citet{Spinks2005}, \citet{Spinks2010}, and \citet{Spinks20140} did not use vouchered specimens we were not able to example the sample individuals. Therefore our specimen classifications were based solely on the geographic information, not explicit assignment in previous studies. Because the exact barriers between different biogeographic regions are unknown and unclear, we represented each hypothesis with two schemes for a total of six different schemes. The schemes differed based on where geographic boundaries were assigned. This changes the classification of certain individuals near the boundaries between groups, providing a test of the robustness of the classification schemes.

@@

Linear discriminate analysis (LDA) attempts to find a linear combination of predictors that best model two or more classes. LDA is very similar to PCA except that instead of finding the linear combination of features that maximize the amount of explained variance in the data, LDA maximizes the differences between classes. The results of this analysis produces a transformation matrix by which the original features can be transformed to reflect the best discrimination between the classes. In this study, we applied LDA to the eigenscores from a subset of the total number of PCs, ranging from two to 6 in increasing order of complexity. In total, this produced nine different LDA scaling matrices.

@@

@@

As part of fitting a random forest model, a ranking of variable importance also is determined. Interestingly, the order of variable importance is not the same as the order of the PCs (\ref{fig:var_imp}). This means that the variance describing the differences between the classes does not align with the major axes of variance (i.e. the PCs). Such a result is expected if variation between classes was extremely fine grained and not a part of the principal form or function of the plastron. Moreover, this result is congruous with the results from the AUC--based model selection for the multinomial logistic regression and LDA models.

@@

@@

The unsupervised learning results indicate only a single group of observations being optimal, which is congruous with the results from the generalizations of the supervised learning models. The classification schemes used in the supervised learning models correspond, loosely, to unsupervised learning solutions with multiple groups. Because unsupervised learning solutions with multiple groups are poor descriptors of the observed variation, it is important to see this generally supported by the supervised learning results.

The results from fitting the various supervised learning models to each of the classification schemes generally shows that no one scheme is ``best.'' Possible explanations include that the genetic differentiation is not associated with plastron shape variation and/or that local selective pressures (e.g. from hydrological regime) overwhelm morphological differentiation. This makes sense given that the plastron is involved in both protection and hydrodynamics, and not necessary mate choice CITATIONS and that shell shape in \textit{E. marmorata} is known to vary among populations inhabiting water bodies with different flow regimes CITATIONS. Plastron shape does not seem to preserve a strong phylogenetic signal at the interspecific level in emydine turtles \citep{Angielczyk2011}, and out current results suggest that this may be the case for phylogeographic signal within emydine species as well. Stated another way, plastron shape may be a poor morphological marker for phylogenetic and phylogeographic studies, despite the ease with which its shape can be quantified. A final possibility (explored below) is that the proposed classification schemes themselves do not represent significant evolutionary lineages.

Both the low AUC values (\(< 0.9\)) and the significant difference between the correctly and incorrectly classified observations support the conclusion that none of the hypothesized classification schemes are good descriptors of the observed plastral variation.

@@

The lack of morphological support for the distinctiveness of \textit{E. pallida} does not, on its own, preclude the recognition of this taxon. However, this apparent lack of congruence does prompt a reexamination of the methods and concepts that led to that taxonomic revision. In other words, before we can assess the significance of the morphological non-diagnosablity, it is essential to evaluate the methods and concepts that led to taxonomic revision.

\citet{Spinks2014} elevated \textit{E. pallida} based on a Bayesian species delimitation analysis of SNP data using BPP \citep{Yang2010b}. However, \citet{Spinks2014} did not heed the caveats about species delimitation methods raised by \citet{Carstens2013}. In addition to specifically addressing the shortcomings of validation methods such as BPP that rely on guide trees and so “should be interpreted with caution,” \citet{Carstens2013} also strongly emphasize that ``Inferences regarding species boundaries based on genetic data alone are likely inadequate, and species delimitation should be conducted with consideration of the life history, geographical distribution, morphology and behaviour (where applicable) of the focal system\dots'' These caveats evoke the development of the Unified Species Concept \citep{Dayrat2005a,DeQueiroz2007b}, Integrative Taxonomy CITATIONS Padial 2010, and other pluralist approaches to species delimitation. None of these considerations were brought to bear on the \textit{E. marmorata} system until now, and in doing so we find the proposal that \textit{E. pallida} is a distinct species to be lacking.

In particular, the natural history and geographical distribution of \textit{E. marmorata} and \textit{E. pallida} make the recognition of these taxa implausible. The data from \citet{Spinks2014} show extensive introgression and admixture in Central California, which is expected because there are no significant barriers to gene flow in this region. Combined with the well-demonstrated ability for testudinoid turtles, including emydids and even \textit{Emys}, to hybridize (e.g. \citealt{Buskirk2005,Spinks2009,Parham2013}) it is hard to imagine how \textit{E. marmorata} and \textit{E. pallida} could maintain their integrity in the face of such admixture. Because the geography, natural history, and demonstrated genetic admixture of \textit{E. marmorata} conflict with the recognition of \textit{E. pallida}, we hypothesize that out inability to classify the morphological data by proposed species is because \textit{E. pallida} is not a distinct species. We agree with \citet{Carstens2013} that ``the inferences drawn from species delimitation studies should be conservative, for in most contexts it is better to fail to delimit species than it is to falsely delimit entities that do not represent actual evolutionary lineages.''

We fully agree with \citet{Spinks2014} that \textit{E. marmorata} (\textit{sensu lato}) is a species deserving of strong conservation efforts, we do not wish to trivialize this need. Moreover, the genetic diversity uncovered by the analysis of \citet{Spinks2014} should be explicitly accounted for in any conservation plan. Given the apparent lack of morphological distinction, however, we consider that this diversity should be considered Evolutionary Significant Units or Distinct Population Segments instead of distinct species.

Finally, it is important to note that the data and analyses we present do not let us definitively say whether the apparent lack of morphological divergence within \textit{E. marmorata} truly reflects the presence of a single species, or if it is an artifact of plastron shape being a poor morphological marker for phylogenetic and phylogeographic divergences. Tests of both our preferred conclusion (\textit{E. marmorata} as a single species) and that \citet{Spinks2014} should include additional tests of species delimitation using different methods and search for corroborating evidence as suggested by \citet{Carstens2013}. From a morphological standpoint, support for the validity of ``\textit{E. Pallida}'' may come from other aspects of morphology, such as carapace shape or other features that might be more informative. Lifewise, further investigation of the phylogeographic utility of plastron shape in other turtle species will help to clarify whether the lack of differentiation seen in \textit{E. marmoarata} is typical or an unusual case.

@@

\begin{table}
  \include{rf_dif}
    %
  \include{mm_dif}
    %
  \include{ll_dif}
  \caption{Results of bootstrap comparisons between the scheme with the highest mean AUC value and all other schemes. An asterix indicates the best scheme. This was done for each of the three modeling techniques included in this study. Probabilities are the percent of comparisons that are greater than the observed difference in means.}
  \label{tab:gen_tests}
\end{table}

\begin{table}
  \begin{scriptsize}
    \include{rf_miss}

    \include{mm_miss}

    \include{ll_miss}
  \end{scriptsize}
  \caption{Results of comparisons between correctly and incorrectly classified observations from the testing data set. For each scheme, the classifications with at least 10 observations were tested. This was done for each of the three modeling techniques included in this study.}
  \label{tab:miss_tests}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/gen_res}
  \caption{Bootstrap distributions for genearlized AUC values for each of the classification schemes. Each row corresponds to a different modeling approach: LDA, LDA using best variables from random forest, multinomial logistic regression, and random forest. Each distribution corresponds to 1000 bootstrap replicates.}
  \label{fig:gen_hist}
\end{figure}

>>>>

\begin{abstract}
  We investigate the morphometric identification of cryptic species using machine-learning approaches by examining their implications for a recently proposed cryptic turtle species (\textit{Emys pallida}). We collected landmark-based morphometric data from 354 adult \textit{E. marmorata/ ``pallida''} museum specimens. We assigned a classification to each specimen for six different binning schemes based on geographic occurrence data recorded in museum collection archives. We used multiple machine learning methods, both unsupervised and supervised, to compare different classification hypotheses and asked whether it is possible to determine which amongst a set of classification hypotheses is best. In addition, we applied the above approach to a “clear-cut” example of seven unambiguously distinct species closely related to \textit{E. marmorata}. The results of this study indicate that there is no clear grouping of \textit{E. marmorata/ ``pallida''} based on plastron shape. In contrast, the analysis of the ``clear-cut'' example demonstrates a near perfect classification, which demonstrates that the methods can recover correct results when an appropriate signal exists. Explanations for the lack of grouping in \textit{E. marmorata} include that possibility that genetic differentiation is not associated with plastron shape variation below the species level and/or that local selective pressures (e.g., from hydrological regime) overwhelm morphological differentiation. A reconsideration of the methods used to delimit \textit{E. ``pallida,''} the lack of barriers to gene flow, the strong evidence for widespread admixture between lineages, and the fact that plastron shape can be used to differentiate other emydine species suggest that its lack of diagnosability most likely reflects the non-distinctiveness of this proposed taxon.
\end{abstract}

@@

% cryptic diversity
%   most species are still deliminated solely based on morphology
Molecular systematics has repeatedly demonstrated the existence of cryptic species that can only be diagnosed using genetic data \citep{Stuart2006,Bickford2007,SchlickSteiner2007,Pfenninger2007,Clare2011,Funk2012}. In attempts to streamline the documentation of biodiversity, several methods of species delimitation that rely almost entirely on genetic data have recently been proposed \citep{Pons2006,Carstens2010,Hausdorf2010,O'Meara2010,Yang2010b,Huelsenbeck2011b}. Although strong caveats on the utility of these methods have been raised \citep{Bauer2000,Carstens2013}, they are nevertheless being used to name species \citep{Leache2010,Spinks2014}.

@@

These considerations have sparked interest in whether geometric morphometric analyses can capture fine-scale variation that can be used for identifying cryptic species. This would make the task of idetifying and maintaining endagered or conserved groups much easier and could contribute to improved classifications of extinct taxa and populations. Most such studies focus on using morphometrics to discover differences between taxa that were identified by other means \citep{Polly2003,Zelditch2004,Gaubert2005b,Gunduz2007,Polly2007a,Demandt2009}. Additionally, there has been work on automated taxon identification and classification of taxa into groups \citep{Baylac2003,Dobigny2003,MacLeod2007,VandenBrink2011}.

@@

% past work on automatic taxon identification and older approaches to classifying taxa
% why use machine learning methods
\subsection{Background and study system}
Machine learning is an extension of known statistical methodology \citep{Hastie2009} that emphasizes high predictive accuracy and generality at the expense of the interpretability of individual parameters. The basic statistical mechanics are supplemented by randomization, sorting, and partitioning algorithms, along with the maximization or minimization of summary statistics, in order to best estimate a general model for all data, both sampled and unsampled \citep{Hastie2009}. Machine learning approaches have found use in medical research, epidemiology, economics and automated image identification such as handwritten zip codes \citep{Hastie2009}. The two major classes of machine learning methods are unsupervised and supervised learning. Unsupervised learning methods are used with unlabeled data where the underlying structure is estimated, and they are analogous to clustering and density estimation methods \citep{Kaufman1990}. Supervised learning methods are used with labeled data where the final output of data is known and the rules for going from input to output are inferred. These are analogous to classification and regression models \citep{Breiman1984}. Our application of the approaches used in this study illustrates only a sampling of the various methods available for clustering observations and fitting classification models.

@@

Here, we used multiple machine learning methods, both unsupervised and supervised, to compare different classification hypotheses. These methods provide different advantages for understanding how to classify taxa, as well as the accuracy of the resulting classifications. Although machine learning methods such as neural networks have been applied to studying shape variation \citep{Baylac2003,Dobigny2003,MacLeod2007,VandenBrink2011}, including in the context of automated taxon identification and classification of groups, the number of cases remains limited. In the current study, we not only consider pure classification accuracy but also use a statistic of classification strength that reflects the rate at which taxa are both accurately and inaccurately classified.

We analyzed the problem of whether there are distinct subspecies or cryptic species within the western pond turtle, \textit{Emys marmorata} \citep{Baird1852} (formerly \emph{Clemmys marmorata}; see \citealp{Feldman2002}). \textit{Emys marmorata} is distributed from northern Washington State, USA to Baja California, Mexico. Traditionally, \textit{E. marmorata} was classified into two named subspecies: the northern \textit{E. marmorata marmorata} and the southern \textit{Emys marmorata pallida} \citep{Seeliger1945}, with a central Californian intergrade zone in between. \textit{Emys marmorata marmorata} is differentiated from \textit{E. marmorata pallida} by the presence of a pair of triangular inguinal scales and darker neck markings. The triangular inguinal plates can sometimes be present in \textit{E. marmorata pallida} although they are considerably smaller. \citet{Seeliger1945} did not formally include the Baja California populations of \textit{E. marmorata} in either taxon, implying the existence of a third distinct but unnamed subspecies.

Previous work on morphological variation in \textit{E. marmorata} has focused primarily on differentiation between populations over a portion of the species' total range \citep{Lubcke2007,Germano2008,Germano2009,Bury2010}; comparatively few studies have included specimens from across the entire range \citep{Holland1992}. Most of these studies considered how local biotic and abiotic factors may contribute to differences in carapace length and found that size can vary greatly between different populations \citep{Lubcke2007,Germano2008,Germano2009}. There also has been interest in size-based sexual dimorphism in \textit{E. marmorata} \citep{Holland1992,Lubcke2007,Germano2009}, with males being on average larger than females based on total carapace length and other linear measurements. However, the quality of size as a classifier of sex can vary greatly between populations \citep{Holland1992} because of the magnitude of size differences among populations \citep{Lubcke2007,Germano2009}. However, the effect of sexual dimorphism on shape, \textit{sensu} \citet{Kendall1977a}, has not been assessed \citep{Holland1992,Lubcke2007,Germano2008}.

@@

More recently, the phylogeography of \textit{E. marmorata} and the possibility of cryptic diversity was investigated using molecular data \citep{Spinks2005,Spinks2010,Spinks2014}. Based on mitochondrial DNA, \citet{Spinks2005} recognized four subclades within \textit{E. marmorata}, a northern clade, a San Joaquin Valley clade, a Santa Barbara clade, and a southern clade. Analyses with nuclear DNA \citep{Spinks2010} and single-nucleotide polymorphism (SNP) data suggest a primarily north--south division in \textit{E. marmorata}, although the dataset differed from that of \citet{Spinks2005} in the location of this break point. These studies discussed the potential taxonomic implications of their results, with \citet{Spinks2014} going so far as to strongly advocate for the recognition of at least two species (\emph{E. marmorata} and \emph{E. pallida}), and a possible third based on populations in Baja California. However, they did not discuss in detail the morphological characters that would help to diagnose these species beyond those specified by \citet{Seeliger1945}. Given that these characters are somewhat variable within the proposed species, and that \citet{Holland1992} described shell shape variation that might be consistent with this taxonomy, a geometric morphometric analysis of shell shape might provide a reliable way to diagnose groups (whether species or subspecies) within \textit{E. marmorata}.

@@

\section{Materials and Methods}
\subsection{Specimens, sampling, morphometrics}
We collected landmark-based morphometric data from 354 adult \textit{E. marmorata} museum specimens. These specimens are a subset of those included in \citet{Angielczyk2007}, \citet{Angielczyk2011}, and \citet{Angielczyk2013a}, essentially representing the adult individuals included in those papers. We chose to focus on adults because significant changes in plastron shape occur over the course of ontogeny in \textit{E. marmorata} and other emydines \citep{Angielczyk2013a}.

We assigned a classification to each specimen for the different binning schemes based on geographic occurrence data recorded in museum collection archives. When precise latitude and longitude information were not available we estimated them from locality information. Because \citet{Spinks2005}, \citet{Spinks2010}, and \citet{Spinks2014} did not use vouchered specimens we were not able to directly sample individuals in their studies. Therefore our specimen classifications were based solely on the geographic information, not explicit assignment using molecular data. Because the exact barriers between different biogeographic regions are unknown and unclear, we represented each hypothesis with two schemes for a total of six different schemes. The schemes differed based on where geographic boundaries were assigned. This changes the classification of certain individuals near the boundaries between groups, providing a test of the robustness of the classification schemes.

@@

Linear discriminate analysis (LDA) attempts to find a linear combination of predictors that best model two or more classes. LDA is very similar to PCA except that instead of finding the linear combination of features that maximize the amount of explained variance in the data, LDA maximizes the differences between classes. The results of this analysis produces a transformation matrix by which the original features can be transformed to reflect the best discrimination between the classes. In this study, we applied LDA to the eigenscores from a subset of the total number of PCs, ranging from two to six in increasing order of complexity. In total, this produced nine different LDA scaling matrices.

@@

\subsection{Comparison with clear-cut example}
In addition to the above analysis of classification schemes of \textit{E. marmorata/pallida}, we applied the above approach to a selection of seven morphologically distinct emydine species. This additional analysis was done in order to confirm the efficacy of our supervised learning approach. The seven species analyzed were \textit{Emys blandigii}, \textit{Terrapene coahuila}, \textit{Clemmys guttata}, \textit{Glyptemys insculpta}, \textit{Glyptemys muhlenbergii}, \textit{Emys orbicularis}, and \textit{Terrapene ornata} with a total of 578 specimens analyzed in total. Again, these data were a subset of those used in \citet{Angielczyk2011} and \citet{Angielczyk2013a}.

As with the \textit{E. marmorata/pallida} analysis, we analyzed plastron shape variation. The support for a seven category classification scheme was evaluated using random forest, linear discriminate analysis, and multinomial logistic regression where the number of PCs used as predictors ranged from two to 11. Data were split into training and testing sets and model performance for both sets was evaluated using the AUC metric.

@@

As part of fitting a random forest model, a ranking of variable importance also is determined. Interestingly, the order of variable importance is not the same as the order of the PCs (\ref{fig:var_imp}). This means that the variance describing the differences between the classes does not align with the major axes of variance (i.e. the PCs). Such a result is expected if variation between classes was extremely fine grained and not a part of the principal form or function of the plastron. Moreover, this result is consistent with the results from the AUC--based model selection for the multinomial logistic regression and LDA models.

@@

\subsection{Comparison with clear-cut example}
In contrast with the above results, the analysis of the seven morphologically-distinct species demonstrates near perfect classification of the both the in-sample and out-of-sample datasets (Table \ref{tab:second_res}). These results demonstrate that when there are strong distinctions between the states of the classification schemes, the methods used here can recover them.

@@

The unsupervised learning results indicate only a single group of observations being optimal, which is consistent with the results from the generalizations of the supervised learning models. The classification schemes used in the supervised learning models correspond, loosely, to unsupervised learning solutions with multiple groups. Because unsupervised learning solutions with multiple groups are poor descriptors of the observed variation, it is important to see this generally supported by the supervised learning results.

The results from fitting the various supervised learning models to each of the classification schemes generally shows that no one scheme is ``best.'' Possible explanations include that the genetic differentiation is not associated with plastron shape variation and/or that local selective pressures (e.g. from hydrological regime) overwhelm morphological differentiation. This makes sense given that the plastron is involved in both protection and hydrodynamics, and not necessary mate choice \citep{Rivera2008,Rivera2011,Stayton2011,Rivera2014} and that shell shape in \textit{E. marmorata} is known to vary among populations inhabiting water bodies with different flow regimes \citep{Holland1992,Lubcke2007,Germano2009}. Plastron shape does not seem to preserve a strong phylogenetic signal at the interspecific level in emydine turtles \citep{Angielczyk2011}, and our current results suggest that this may be the case for phylogeographic signal within emydine species as well. A final possibility (explored below) is that the proposed classification schemes themselves do not represent significant evolutionary lineages.

Both the low AUC values (\(< 0.9\)) and the significant difference between the correctly and incorrectly classified observations support the conclusion that none of the hypothesized classification schemes are good descriptors of the observed plastral variation within \textit{E. marmorata}.

Nevertheless, it is important to note that plastron shape is an extremely effective method for differentiating members of the other seven species we investigated. The magnitude of shape differences between the species (measured as Procrustes distance between species' mean shapes) is approximately an order of magnitude greater than the differences between the \textit{E. marmorata} subgroups, and the machine learning methods had no trouble accurately classifying the specimens correctly. These results demonstrate that plastron shape is normally a good marker for species delineation iin the closest relatives of \textit{E. marmorata}, and that our lack of results for \textit{E. marmorata} is not simply a shortcoming of the methods we applied. Indeed, they beg the question of what factors have suppressed morphological differentiation of plastron shape in \textit{E. marmorata} and \textit{E. pallida} if they are distinct species. Invoking issues such as the role of the plastron in protection or the need for streamlining are insufficient because the other species are expected to be subject to similar constraints \citep{Stayton2011,Pollyb}.

@@

The lack of morphological support for the distinctiveness of \textit{E. pallida} does not, on its own, preclude the recognition of this taxon. However, this apparent lack of congruence does prompt a reexamination of the methods and concepts that led to that taxonomic revision, especially considering that plastron shape is normally a strong indicator of species differences amoung emydines. In other words, before we can assess the significance of the morphological non-diagnosablity, it is essential to evaluate the methods and concepts that led to the initial taxonomic revision.

\citet{Spinks2014} elevated \textit{E. pallida} based on a Bayesian species delimitation analysis of SNP data using BPP \citep{Yang2010b}. However, \citet{Spinks2014} did not heed the caveats about species delimitation methods raised by \citet{Carstens2013}. In addition to specifically addressing the shortcomings of validation methods such as BPP that rely on guide trees and ``should be interpreted with caution,'' \citet{Carstens2013} also strongly emphasize that ``Inferences regarding species boundaries based on genetic data alone are likely inadequate, and species delimitation should be conducted with consideration of the life history, geographical distribution, morphology and behaviour (where applicable) of the focal system\dots'' These caveats evoke the development of the Unified Species Concept \citep{Dayrat2005a,DeQueiroz2007b}, Integrative Taxonomy \citep{Padial2010}, and other pluralist approaches to species delimitation. None of these considerations were brought to bear on the \textit{E. marmorata} system until now, and in doing so we find the proposal that \textit{E. pallida} is a distinct species to be lacking in a normally robust morphological marker.

The natural history and geographical distribution of \textit{E. marmorata} and \textit{E. pallida} also make the recognition of these taxa implausible. The data from \citet{Spinks2014} show extensive introgression and admixture in Central California, which is expected because there are no significant barriers to gene flow in this region. Combined with the well-demonstrated ability for testudinoid turtles, including emydids and even \textit{Emys}, to hybridize (e.g. \citealt{Buskirk2005,Spinks2009,Parham2013}) it is hard to imagine how \textit{E. marmorata} and \textit{E. pallida} could maintain their integrity in the face of such admixture. Because the geography, natural history, demonstrated genetic admixture of \textit{E. marmorata}, and comparisons with other morphologically-diagnosable species conflict with the recognition of \textit{E. pallida}, we hypothesize that out inability to classify the morphological data by proposed species is because \textit{E. pallida} is not a distinct species. We agree with \citet{Carstens2013} that ``the inferences drawn from species delimitation studies should be conservative, for in most contexts it is better to fail to delimit species than it is to falsely delimit entities that do not represent actual evolutionary lineages.''

We fully agree with \citet{Spinks2014} that \textit{E. marmorata} (\textit{sensu lato}) is a species deserving of strong conservation efforts, and we do not wish to trivialize this need. Moreover, the genetic diversity uncovered by the analysis of \citet{Spinks2014} should be explicitly accounted for in any conservation plan. Given the apparent lack of morphological distinction, however, we consider that this diversity should be considered Evolutionary Significant Units or Distinct Population Segments instead of distinct species.

Finally, it is important to note that the data and analyses we present do not let us definitively say whether the apparent lack of morphological divergence within \textit{E. marmorata} truly reflects the presence of a single species, or if it is an artifact of plastron shape being a poor morphological marker for phylogenetic and phylogeographic divergences, because we could not carry out our morphometric analyses on the specimens from which the genetic data were obtained. The comparisons with the other emydine species, suggests it is because \textit{E. marmorata} is a single species. However, tests of both our preferred conclusion (\textit{E. marmorata} as a single species) and that \citet{Spinks2014} should include morphological and molecular analyses of the same set of voucher specimens, as well as additional tests of species delimitation using alternative methods and corroborating evidence as suggested by \citet{Carstens2013}. From a morphological standpoint, support for the validity of ``\textit{E. Pallida}'' may come from other aspects of morphology, such as carapace shape or other features. Likewise, further investigation of the phylogeographic utility of plastron shape in other turtle species will help to clarify whether the lack of differentiation seen in \textit{E. marmoarata}, and the strong differentiation among the other emydines, is typical or an unusual case.

@@

\begin{table}
  \centering
  \begin{tabular}{ l r r r }
    \hline
    Scheme & RF & MLR & LDA \\
    \hline
    \hline
    Morph 1 & \(\ast\) & \(\ast\) & 1 \\
    Morph 2 & 0.79 & 0.55 & 1 \\
    Mito 1 & 0.89 & 0.94 & 1 \\ 
    Mito 2 & 0.82 & 0.57 & \(\ast\) \\ 
    Mito 3 & 0.79 & 0.69 & 0.73 \\ 
    Nuclear & 0.79 & 0.96 & 0.96 \\ 
    \hline
  \end{tabular}
  \caption{Results of bootstrap comparisons between the scheme with the highest mean AUC value and all other schemes, defined P(best - other \(>\) 0). An asterix indicates the best scheme. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA). Probabilities are the percent of comparisons that are greater than the observed difference in means.}
  \label{tab:gen_tests}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{ l | l r r r }
    \hline
    Scheme & Class & RF & MLR & LDA \\
    \hline
    \hline
    \multirow{2}{*}{Morph 1} & CCR & 0.77 & 1.00 & 1.00 \\ 
    & marm & 0.87 & 0.93 & 1.00 \\ 
    \hline
    \multirow{2}{*}{Morph 2} & CCR & 0.88 & 1.00 & 1.00 \\ 
    & marm & 1.00 & 1.00 & 0.98 \\ 
    \hline
    \multirow{2}{*}{Mito 1} & CCR & 0.94 & 0.99 & 0.98 \\ 
    & marm & 0.99 & 0.98 & 0.47 \\ 
    \hline
    \multirow{2}{*}{Mito 2} & marm & 0.85 & 0.97 & 0.96 \\ 
    & pall & 0.94 & 1.00 & 1.00 \\ 
    \hline
    \multirow{2}{*}{Mito 3} & 1 & 0.40 & 0.92 & 0.90 \\ 
    & 3 & 0.97 & 0.99 & 1.00 \\ 
    \hline
    \multirow{2}{*}{Nuclear} & marm & 1.00 & 1.00 & 1.00 \\ 
    & pall & 0.99 & 1.00 & 1.00 \\ 
    \hline
  \end{tabular}
  \caption{Results of comparisons between correctly and incorrectly classified observations from the testing data set, defined P(distance - simulated \(>\) 0). For each scheme, the classifications with at least 10 observations were tested. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA).}
  \label{tab:miss_tests}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[height = \textheight, width = \textwidth, keepaspectratio = true]{figure/gen_res}
  \caption{Bootstrap distributions for genearlized AUC values for each of the classification schemes. Each row corresponds to a different modeling approach: multinomial logistic regression (MLR), random forest (RF), linear discriminate analysis (LDA), and LDA using best variables from random forest. Each distribution corresponds to 1000 bootstrap replicates.}
  \label{fig:gen_hist}
\end{figure}

\begin{table}[ht]
  \centering
  \begin{tabular}{ l r r r }
    \hline
    & \# predictors & In-sample AUC & Out-of-sample AUC \\ 
    \hline
    RF &   11 & 1.000 & 0.999 \\ 
    MLR &    6 & 1.000 & 0.998 \\ 
    LDA &   10 & 1.000 & 1.000 \\ 
    \hline
  \end{tabular}
  \caption{Results from classification model estimates of the secondary, multi-species dataset. Models are random forest (RF), linear discriminate analysis (LDA), and multinomial logistic regression (MLR). Comparison of in-sample and out-of-sample AUC of the best performing model, along with the number of predictors. AUC values range between 0.5 (no better than random) and 1 (perfect classification).}
  \label{tab:second_res}
\end{table}
